\chapter*{Acknowledgements}

More than just being the result of a year of research, this thesis marks the end of a twenty-six year winding road through the education system, ten of which have been spent at this university. I have had a lot to learn besides just physics, and I could not be prouder of how far I have come; this document stands as a testament to that. It is my great pleasure to use this space to extend gratitude to all those who have been with me through this phase of my life.

First of all, a name on the front cover does not begin to do justice to the positive impact that Cristiane and Marcus have had on my personal growth and well-being over the past year. I could not have asked for a more understanding and encouraging duo of supervisors. Cris, you went to great lengths this year to invest additional time and energy in me, despite having so little time to spare. Your uncompromising faith in the ability of others is an inspiration, and it has had a contagious effect on me. I never felt like you doubted me for a moment, even when I had nothing to show for it---and look how I proved you right. 

Marcus, I may not have been the most straightforward first student to supervise, but I have a lot of respect for the way you handled things. Your willingness to normalise and prioritise mental health over work when needed has been a great encouragement; this year might have looked very different for me if I had not felt comfortable opening up to you in the first months. In the moments when we did get around to discussing the content, you always had insightful comments, and I think back fondly to us geeking out over cool topology together.

The rest of the research group has also been like a warm blanket (we still use the one we won at Sinterklaas!). I may not have interacted with all of you as much as I would have liked, but I really appreciated the sense of friendship and support in the group. I especially thank Lumen and Anouar for some insightful discussions, and Robin for being a great example and taking the time to chat about neurodivergence.

Outside of the studies, more people have impacted my life than I can reasonably find the space to thank here. The family and in-laws who never stopped believing in me; the friends who lent me their ears, kept me close to Christ and distracted me with board games; my therapist André and all those in my support group who have encouraged me to grow---you all know who you are, and I could not have done this without you.

Finally, my greatest thanks go out to my wife and favourite human being \mbox{Annemarijn}, to whom it is my absolute pleasure to dedicate this thesis. Your love and support have been unwavering, even through the hardest years of my life, and words can never adequately express how much I appreciate you.